---
title: "Simulation Results Summary"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(here))

theme_set(theme_minimal() + 
            theme(legend.position = "bottom",
                  ))
# Define a custom color palette for the methods
my_colors <- c(
  "BKMR" = "#440154",
  "BWS" = "#414487",
  "qgcomp" = "#2A788E",
  "qgcomp nonlinear" = "#7AD151",
  "WQS" = "#FDE725",
  "2iWQS" = "#E03C8A",
  "2iWQS:pos" = "#B481BB"
)
fig_dir <- "/Users/weijia/Research/Frailty/analysis/figures"
```

## Load simulation outputs

```{r}
folder_path <- here("results", "k5")

rda_files <- list.files(folder_path, pattern = "\\.RDA$", full.names = TRUE)
length(rda_files)

# Robust loader: guarantees we extract object named `results`
all_results <- vector("list", length(rda_files))
for (i in seq_along(rda_files)) {
  e <- new.env()
  load(rda_files[[i]], envir = e)
  stopifnot(exists("results", envir = e))
  all_results[[i]] <- get("results", envir = e)
}

folder_path <- here("results", "qgcomp")

rda_files <- list.files(folder_path, pattern = "\\.RDA$", full.names = TRUE)
length(rda_files)

# Robust loader: guarantees we extract object named `results`
all_qgcomp <- vector("list", length(rda_files))
for (i in seq_along(rda_files)) {
  e <- new.env()
  load(rda_files[[i]], envir = e)
  stopifnot(exists("results", envir = e))
  all_qgcomp[[i]] <- get("results", envir = e)
}
```

## Flatten outputs to tables

```{r}
all_coef <- list()
all_weights <- list()
all_bkmr <- list()

for (batch_res in all_results) {
  for (iter_res in batch_res) {
    if (is.null(iter_res)) next
    if (!is.null(iter_res$coef))    all_coef[[length(all_coef) + 1]] <- iter_res$coef
    if (!is.null(iter_res$weights)) all_weights[[length(all_weights) + 1]] <- iter_res$weights
    if (!is.null(iter_res$bkmr))    all_bkmr[[length(all_bkmr) + 1]] <- iter_res$bkmr
  }
}

all_coef_qg <- list()
all_weights_qg <- list()

for (batch_res in all_qgcomp) {
  for (iter_res in batch_res) {
    if (is.null(iter_res)) next
    if (!is.null(iter_res$coef))    all_coef_qg[[length(all_coef_qg) + 1]] <- iter_res$coef
    if (!is.null(iter_res$weights)) all_weights_qg[[length(all_weights_qg) + 1]] <- iter_res$weights
  }
}
  
weights_df <- bind_rows(all_weights) %>%
  filter(!(method %in% c("qgcomp.noboot", "qgcomp.boot"))) %>%
  bind_rows(all_weights_qg) %>% 
  distinct() %>%
  mutate(method = case_when(str_detect(method, "BKMR") ~ "BKMR",
                             str_detect(method, "BWS") ~ "BWS",
                             str_detect(method, "qgcomp.noboot") ~ "qgcomp",
                             str_detect(method, "qgcomp.boot") ~ "qgcomp nonlinear",
                             str_detect(method, "WQS2: pos") ~ "2iWQS:pos",
                             str_detect(method, "WQS2: neg") ~ "2iWQS:neg",
                             .default = "WQS"),
         method = factor(method, levels = c("BKMR", "BWS", "qgcomp", "qgcomp nonlinear", "WQS", "2iWQS", "2iWQS:pos", "2iWQS:neg")),
         scenario = factor(scenario, levels = c("null", "single", "homogeneous", "heterogeneous", "nonlinear", "interactive")))

bkmr_df <- bind_rows(all_bkmr) %>% distinct()
# --- BKMR: already returns Delta(q, 0.5) for q=0.75 (empirical quantiles) ---
bkmr_delta <- bkmr_df %>%
  filter(abs(quantile - 0.75) < 1e-12) %>%
  mutate(
    method = "BKMR",
    term = "delta",
    estimate = est,
    lb = est - 1.96 * sd,
    ub = est + 1.96 * sd
  ) %>%
  select(batch, scenario, n, p, rho_X, sigma, seed, method, term, estimate, lb, ub)

coef_df <- bind_rows(all_coef) %>%
  filter(!(method %in% c("qgcomp.noboot", "qgcomp.boot"))) %>%
  bind_rows(all_coef_qg) %>% 
  bind_rows(bkmr_delta) %>%
  distinct() %>%
  mutate(method = case_when(str_detect(method, "BKMR") ~ "BKMR",
                             str_detect(method, "BWS") ~ "BWS",
                             str_detect(method, "qgcomp.noboot") ~ "qgcomp",
                             str_detect(method, "qgcomp.boot") ~ "qgcomp nonlinear",
                             str_detect(method, "WQS2: pos") ~ "2iWQS:pos",
                             str_detect(method, "WQS2: neg") ~ "2iWQS:neg",
                             .default = "WQS"),
         method = factor(method, levels = c("BKMR", "BWS", "qgcomp", "qgcomp nonlinear", "WQS", "2iWQS", "2iWQS:pos", "2iWQS:neg")),
         scenario = factor(scenario, levels = c("null", "single", "homogeneous", "heterogeneous", "nonlinear", "interactive")))

```

## Define the true joint-effect contrast Δ(75,50)

We use a common, method-agnostic contrast:
\[
\Delta(75,50) = E[Y \mid X_1=\cdots=X_p=q_{0.75}] - E[Y \mid X_1=\cdots=X_p=q_{0.50}]
\]
where \(q_{0.50}=0\), \(q_{0.75}=\Phi^{-1}(0.75)\).

```{r}
true_joint_effect <- function(scenario, p = 5) {
  q <- qnorm(0.75)
  X75 <- rep(q, p)
  X50 <- rep(0, p)

  f <- function(X, scenario) {
    if (scenario == "null") return(0)
    if (scenario == "single") return(0.25 * X[1])
    if (scenario == "homogeneous") return(sum(rep(0.0625, 4) * X[1:4]))
    if (scenario == "heterogeneous") return(0.25 * X[1] - 0.15 * X[2])
    if (scenario == "nonlinear") {
      # MUST match simulate_data.R
      return(0.35 * X[1] - 0.15 * X[1]^2 + 0.20 * sin(pi * X[2]))
    }
    if (scenario == "interactive") {
      I1 <- as.numeric(X[1] >= q)
      I2 <- as.numeric(X[2] >= q)
      return(0.25 * I1 + 0.15 * I2 - 0.15 * I1 * I2)
    }
    stop("Unknown scenario")
  }

  f(X75, scenario) - f(X50, scenario)
}

truth_tbl <- expand_grid(
  scenario = unique(coef_df$scenario),
  p = unique(coef_df$p)
) %>%
  mutate(true_delta = map2_dbl(scenario, p, true_joint_effect))

truth_tbl
```

## Construct method-specific estimated Δ(75,50)

For most methods, we use their reported joint-effect parameter as the estimate of the contrast (consistent with your previous summaries). For BKMR, we use the OverallRiskSummaries at quantile = 0.75 (already relative to q.fixed=0.5).

```{r}
# Continuous contrast step for N(0,1): median=0, 75th=qnorm(0.75)
step_cont <- qnorm(0.75) - qnorm(0.5)  # = qnorm(0.75)

# --- WQS (single index): use its reported slope as the contrast ---
wqs_delta <- coef_df %>%
  filter(method == "WQS") %>%
  # keep the mixture term only (usually "wqs")
  arrange(batch, seed) %>%
  select(batch, scenario, n, p, rho_X, sigma, seed, method, term, estimate, lb, ub) %>%
  select(-term)

# --- WQS2 (two indices): continuous contrast corresponds to +1 quartile-category shift for both indices
# so the joint contrast is beta_pos + beta_neg
wqs2_delta <- coef_df %>%
  filter(method %in% c("2iWQS:pos", "2iWQS:neg")) %>%
  filter(!term %in% c("(Intercept)", "Intercept")) %>%
  mutate(component = ifelse(grepl("pos", method), "pos", "neg")) %>%
  select(batch, scenario, n, p, rho_X, sigma, seed,
         component, estimate, lb, ub) %>%
  tidyr::pivot_wider(
    names_from = component,
    values_from = c(estimate, lb, ub),
    values_fn = dplyr::first
  ) %>%
  mutate(
    method = "2iWQS",
    estimate = estimate_pos + estimate_neg,
    lb = lb_pos + lb_neg,
    ub = ub_pos + ub_neg
  ) %>%
  select(batch, scenario, n, p, rho_X, sigma, seed, method, estimate, lb, ub)

# --- qgcomp.noboot (degree=1): psi1 corresponds to a 1-category quartile increase -> matches median->75th
qgcomp_delta <- coef_df %>%
  filter(method == "qgcomp") %>%
  select(batch, scenario, n, p, rho_X, sigma, seed, method, term, estimate, lb, ub) %>%
  filter(term != "(Intercept)") %>%
  select(-term) %>%
  mutate(
    method = "qgcomp")

# --- qgcomp.boot nonlinear (degree=2): contrast for 1-step increase is psi1 + psi2
# (Intercept), psi1, psi2 exist in your updated model
qgcomp_nonlinear_delta <- coef_df %>%
  filter(method == "qgcomp nonlinear") %>%
  select(batch, scenario, n, p, rho_X, sigma, seed, method, term, estimate, lb, ub) %>%
  tidyr::pivot_wider(names_from = term, values_from = c(estimate, lb, ub)) %>%
  mutate(
    estimate = estimate_psi1 + estimate_psi2,
    lb = lb_psi1 + lb_psi2,
    ub = ub_psi1 + ub_psi2
  ) %>%
  select(batch, scenario, n, p, rho_X, sigma, seed, method, estimate, lb, ub)

# --- BWS: theta1 is slope per 1-unit increase in weighted index sum(w_j X_j).
# Under continuous contrast where all X_j shift by step_cont and sum(w)=1:
# Delta = theta1 * step_cont
bws_delta <- coef_df %>%
  filter(method == "BWS") %>%
  select(batch, scenario, n, p, rho_X, sigma, seed, method, term, estimate, lb, ub) %>%
  filter(term == "theta1") %>%
  mutate(
    estimate = estimate * step_cont,
    lb = lb * step_cont,
    ub = ub * step_cont
  ) %>%
  select(-term)

# --- Combine all methods ---
delta_df <- bind_rows(
  bkmr_delta,
  wqs_delta,
  wqs2_delta,
  qgcomp_delta,
  qgcomp_nonlinear_delta,
  bws_delta
) %>%
  left_join(truth_tbl, by = c("scenario", "p")) %>%
  mutate(
    method = factor(method, levels = c("BKMR", "BWS", "qgcomp", "qgcomp nonlinear", "WQS", "2iWQS", "2iWQS:pos", "2iWQS:neg")),
         scenario = factor(scenario, levels = c("null", "single", "homogeneous", "heterogeneous", "nonlinear", "interactive")),
    bias = estimate - true_delta,
    sq_error = (estimate - true_delta)^2,
    covered = ifelse(!is.na(lb) & !is.na(ub),
                     (lb <= true_delta & true_delta <= ub),
                     NA)
  )

```

## Type I error and power (by method, scenario, rho, sigma)

We use CI inclusion logic:
- Type I error: null scenario, reject if CI excludes 0.
- Power: non-null scenarios, reject if CI excludes 0.

```{r}
df_type1 <- coef_df %>%
  filter(scenario == "null" & method != "qgcomp nonlinear") %>%
  mutate(reject0 = (lb > 0 | ub < 0)) %>%
  group_by(method, scenario, rho_X, sigma) %>%
  summarise(type1 = round(mean(reject0, na.rm = TRUE),2))
df_type1
```


```{r}
df_power <- coef_df %>%
  filter(scenario != "null" & !(method %in% c("2iWQS:neg", "qgcomp nonlinear")) & term != "(intercept)") %>%
  mutate(reject0 = (lb > 0 | ub < 0)) %>%
  group_by(scenario, rho_X, sigma, method) %>%
  summarise(se = sd(reject0, na.rm = TRUE)/sqrt(500),
            tcrit  = qt(0.975, df = 500 - 1),   # 97.5% quantile of t_{n-1}
            mean = round(mean(reject0, na.rm = TRUE),2),
            lower  = round(mean - tcrit * se,2),                            # lower 95% CI
            upper  = round(mean + tcrit * se,2))
  # pivot_wider(names_from = method, values_from = mean)

# Bar plot: power
plot_power <- df_power %>%
  filter(sigma == 1) %>%
  ggplot(aes(factor(method), mean, fill = method)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = my_colors) +
  # geom_hline(yintercept = 0.95, color = "red", lty = "dashed")+
  facet_grid(rho_X~scenario, labeller = labeller(rho_X = label_both, scenario = label_value), scales = "free") +
  labs(x = "", y = "Mean Power", fill = "Method") +
  theme(
    axis.title.x=element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
  )
plot_power
```

```{r}
power_plot_df <- power_tbl %>%
  mutate(
    rho_X = factor(rho_X),
    scenario = factor(scenario),
    method = factor(method)
  )

ggplot(power_plot_df, aes(x = sigma, y = power, color = method, group = method)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  facet_grid(rho_X~scenario) +
  labs(
    x = "Number of exposures (K)",
    y = "Power",
    color = "Method",
    title = "Power vs K by scenario and correlation"
  ) +
  theme(
    legend.position = "top",
    strip.text = element_text(face = "bold")
  )
```


## Bias, RMSE, and 95% coverage

```{r}
perf_tbl <- delta_df %>%
  filter(scenario != "null") %>%
  group_by(method, scenario, rho_X, sigma) %>%
  summarise(
    bias = mean(bias, na.rm = TRUE),
    rmse = sqrt(mean(sq_error, na.rm = TRUE)),
    coverage = mean(covered, na.rm = TRUE),
    .groups = "drop"
  )

perf_tbl

plot_coverage <- perf_tbl %>%
  filter(sigma == 1) %>%
  ggplot(aes(factor(method), coverage, fill = method)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = my_colors) +
  # geom_hline(yintercept = 0.95, color = "red", lty = "dashed")+
  facet_grid(rho_X~scenario, labeller = labeller(rho_X = label_both, scenario = label_value), scales = "free") +
  labs(x = "", y = "Mean 95% CI Coverage", fill = "Method") +
  theme(
    axis.title.x=element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
  )
plot_coverage
```

## Weight sensitivity/specificity

We normalize qgcomp directional weights within each replicate so selection thresholds are meaningful.

Assume the “true active” set by scenario:
- single: X1
- homogeneous: X1–X4
- heterogeneous: X1 and X2 (both active; sign ignored for selection)
- nonlinear: X1 and X2
- interactive: X1 and X2

```{r}
true_active <- function(scenario) {
  if (scenario == "single") return(c("X1"))
  if (scenario == "homogeneous") return(paste0("X", 1:4))
  if (scenario == "heterogeneous") return(c("X1", "X2"))
  if (scenario == "nonlinear") return(c("X1", "X2"))
  if (scenario == "interactive") return(c("X1", "X2"))
  if (scenario == "null") return(character(0))
  stop("Unknown scenario")
}

# normalize weights for qgcomp to sum abs=1 per replicate-method
# weights_norm <- weights_df %>%
#   group_by(batch, seed, method) %>%
#   mutate(
#     abs_sum = sum(abs(estimate), na.rm = TRUE),
#     estimate_norm = ifelse(method == "qgcomp.noboot" & abs_sum > 0,
#                            abs(estimate) / abs_sum,
#                            estimate)
#   ) %>%
#   ungroup()

# selection rule:
# - For WQS/BWS (weights sum to 1): select if weight > 1/p
# - For qgcomp.noboot: select if normalized abs(weight) > 1/p
# - For BKMR PIP: select if PIP > 0.5 (conservative)
weights_sel <- weights_df %>%
  mutate(
    thresh = case_when(
      method == "BKMR" ~ 0.5,
      # method == "qgcomp.noboot" ~ 1 / p,
      TRUE ~ 1 / p
    ),
    selected = case_when(
      method == "BKMR" ~ (estimate > thresh),
      method == "qgcomp" ~ (abs(estimate) > thresh),
      TRUE ~ (estimate > thresh)
    )
  )

sens_spec <- weights_sel %>%
  filter(scenario != "null") %>%
  group_by(method, scenario, rho_X, sigma, batch, seed) %>%
  summarise(
    active = list(true_active(first(scenario))),
    sens = {
      a <- active[[1]]
      if (length(a) == 0) NA_real_ else mean(selected[mix_name %in% a], na.rm = TRUE)
    },
    spec = {
      a <- active[[1]]
      inact <- setdiff(unique(mix_name), a)
      if (length(inact) == 0) NA_real_ else mean(!selected[mix_name %in% inact], na.rm = TRUE)
    },
    .groups = "drop"
  ) %>%
  group_by(method, scenario, rho_X, sigma) %>%
  summarise(
    sensitivity = mean(sens, na.rm = TRUE),
    specificity = mean(spec, na.rm = TRUE),
    .groups = "drop"
  )

sens_spec

plot_se <- sens_spec %>%
  filter(sigma == 1) %>%
  ggplot(aes(factor(method), sensitivity, fill = method)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = my_colors) +
  # geom_hline(yintercept = 0.95, color = "red", lty = "dashed")+
  facet_grid(rho_X~scenario, labeller = labeller(rho_X = label_both, scenario = label_value), scales = "free") +
  labs(x = "", y = "Mean Sensitivity", fill = "Method") +
  theme(
    axis.title.x=element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
  )
plot_se

plot_sp <- sens_spec %>%
  filter(sigma == 1) %>%
  ggplot(aes(factor(method), specificity, fill = method)) + 
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = my_colors) +
  # geom_hline(yintercept = 0.95, color = "red", lty = "dashed")+
  facet_grid(rho_X~scenario, labeller = labeller(rho_X = label_both, scenario = label_value), scales = "free") +
  labs(x = "", y = "Mean Specificity", fill = "Method") +
  theme(
    axis.title.x=element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
  )
plot_sp
```
