---
title: "Simulation Results"
author: "Weijia Qian"
date: "2024-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(CCMHr)
library(tidyverse)
library(knitr)

# load simulation results
folder_path <- "/Users/weijia/Research/Huels/results/20250513"
rda_files <- list.files(path = folder_path, pattern = "\\.RDA$", full.names = TRUE)
all_results <- list()
for (i in 1:length(rda_files)) {
  all_results[[i]] <- loadRDa(rda_files[[i]])
}
```

### Updates
- Ran simulations with 100 iterations and analyzed the results
- In 2 high-correlation scenarios, gWQS threw an error:
Error in vif.default(i$mfit$m_f) : 
  there are aliased coefficients in the model
Calls: gwqs ... resolve.list -> signalConditionsASAP -> signalConditions
Execution halted
3. For BKMR, I used the significance of the relative overall risk at 75th percentile (compared to 50 percentile) to calculate type1 error and power.

### Type 1 Error
```{r}
all_coef_list <- list() 

for (scenario in all_results) {
  for (iteration in scenario) {
    all_coef_list[[length(all_coef_list) + 1]] <- iteration$coef
  }
}

all_coef <- bind_rows(all_coef_list)

all_bkmr <- data.frame()
for (scenario in all_results) {
  for (iteration in scenario) {
    all_bkmr <- rbind(all_bkmr, iteration$bkmr)
  }
}

df_bkmr_q3 <- all_bkmr %>%
  filter(quantile == 0.75) %>%
  mutate(method = "BKMR",
         estimate = est,
         lb = est - 1.96 * sd,
         ub = est + 1.96 * sd) %>%
  select(-quantile, -est, -sd)

all_coef <- bind_rows(all_coef, df_bkmr_q3)

all_coef <- all_coef %>%
  mutate(scenario = case_when(scenario %in% c(1,2,3) ~ "Null",
                              scenario %in% c(4,5,6) ~ "Single effect",
                              scenario %in% c(7,8,9) ~ "Homogeneous effect",
                              scenario %in% c(10,11,12) ~ "Heterogeneous effect",
                              scenario %in% c(13,14,15) ~ "Confounding",
                              scenario %in% c(16,17,18) ~ "Nonlinear effect",
                              scenario %in% c(19,20,21) ~ "Nonadditive effect"),
         param = case_when(grepl("^psi2", rownames(all_coef)) ~ "psi2",
                           grepl("^psi3", rownames(all_coef)) ~ "psi3",
                           .default = "psi1"),
         is_significant = (lb > 0 | ub < 0)) %>%
  filter(seed != 664)
all_coef$scenario <- factor(all_coef$scenario, 
                            levels = c("Null", "Single effect", "Homogeneous effect", "Heterogeneous effect", "Confounding", 
                                          "Nonlinear effect", "Nonadditive effect"))

### Type 1 error
df_type1 <- all_coef %>%
  filter(scenario == "Null") %>%
  group_by(scenario, rho_X, method) %>%
  summarise(type1_error = mean(is_significant, na.rm = TRUE)) %>%
  pivot_wider(names_from = method, values_from = type1_error)
kable(df_type1, format = "markdown", caption = "Type 1 error by scenario and method")

df_type1_neg <- all_coef %>%
  filter(method == "WQS2: neg" & scenario != "Heterogeneous effect") %>%
  group_by(scenario, rho_X, method) %>%
  summarise(type1_error = mean(is_significant, na.rm = TRUE)) %>%
  pivot_wider(names_from = rho_X, values_from = type1_error)
kable(df_type1_neg, format = "markdown", caption = "Mean type 1 error of 2iWQS negative index")
```

### Power
- BKMR and linear qgcomp model have poor power in the heterogeneous effect scenario.
- Power of 2iWQS decreases significantly when exposures are highly correlated.

```{r}
### Power
df_power <- all_coef %>%
  filter(scenario != "Null" & !(method %in% c("WQS2: neg", "qgcomp.boot"))) %>%
  group_by(scenario, rho_X, method) %>%
  summarise(power = mean(is_significant, na.rm = TRUE)) %>%
  pivot_wider(names_from = method, values_from = power)
kable(df_power, format = "markdown", digits = 2, caption = "Mean power by scenario and method")

# ggplot(df_power, 
#        aes(x = method, y = power, fill = method)) +
#   geom_bar(stat = "identity", position = "stack") +
#   facet_wrap(~ scenario + rho_X) +                           
#   labs(x = "Method", y = "Power", title = "Mean power by scenario and method") +
#   theme_minimal() +
#   theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

df_power_2 <- all_coef %>%
  filter(method == "qgcomp.boot") %>%
  group_by(scenario, rho_X, method, param) %>%
  summarise(power = mean(is_significant, na.rm = TRUE)) %>%
  pivot_wider(names_from = param, values_from = power)
kable(df_power_2, format = "markdown", caption = "Mean power of nonlinear qgcomp model")
```

### Sensitivity and Specificity
- BKMR has the highest sensitivity across all scenarios.
- In the heterogeneous effect scenario, nonlinear qgcomp model has higher sensitivity than BWS and WQS.
- BWS has the highest specificity across all scenarios.
- BKMR has the lowest specificity in many scenarios.
```{r}
all_weights <- data.frame()
for (scenario in all_results) {
  for (iteration in scenario) {
    all_weights <- rbind(all_weights, iteration$weights)
  }
}

all_weights <- all_weights %>%
  filter(method != "WQS2: neg") %>%
  mutate(scenario = case_when(scenario %in% c(1,2,3) ~ "Null",
                              scenario %in% c(4,5,6) ~ "Single effect",
                              scenario %in% c(7,8,9) ~ "Homogeneous effect",
                              scenario %in% c(10,11,12) ~ "Heterogeneous effect",
                              scenario %in% c(13,14,15) ~ "Confounding",
                              scenario %in% c(16,17,18) ~ "Nonlinear effect",
                              scenario %in% c(19,20,21) ~ "Nonadditive effect"),
         is_important = ifelse(method == "BKMR", estimate > 0.5, abs(estimate) > 0.1),
         true_predictor = ifelse((mix_name == "X1" & scenario != "Null")|
                                 (mix_name == "X2" & scenario %in% c("Homogeneous effect", "Nonadditive effect")) |
                                 (mix_name == "X2" & scenario == "Heterogeneous effect" & method != "WQS2: pos") |
                                 (mix_name %in% c("X3","X4") & scenario == "Homogeneous effect"), TRUE, FALSE),
         true_weight = case_when(scenario == "Null" ~ NA,
                                 scenario %in% c("Single effect", "Confounding", "Nonlinear effect") & mix_name == "X1" ~ 1,
                                 scenario %in% c("Single effect", "Confounding", "Nonlinear effect") & mix_name != "X1" ~ 0,
                                 scenario == "Homogeneous effect" & mix_name %in% c("X1","X2","X3","X4") ~ 0.25,
                                 scenario == "Homogeneous effect" & !(mix_name %in% c("X1","X2","X3","X4")) ~ 0,
                                 scenario == "Heterogeneous effect" & mix_name == "X1" & method != "WQS2: pos" ~ 0.625,
                                 scenario == "Heterogeneous effect" & mix_name == "X1" & method == "WQS2: pos" ~ 1,
                                 scenario == "Heterogeneous effect" & mix_name == "X2" & method != "WQS2: pos" ~ 0.375,
                                 scenario == "Heterogeneous effect" & mix_name == "X2" & method == "WQS2: pos" ~ 0,
                                 scenario == "Heterogeneous effect" & mix_name %in% c("X3","X4","X5") ~ 0,
                                 scenario == "Nonadditive effect" & mix_name == "X1" ~ 0.625,
                                 scenario == "Nonadditive effect" & mix_name == "X2" ~ 0.375,
                                 scenario == "Nonadditive effect" & mix_name %in% c("X3","X4","X5") ~ 0),
         method = ifelse(method == "qgcomp.noboot", "qgcomp", method))
all_weights$scenario <- factor(all_weights$scenario, 
                               levels = c("Null", "Single effect", "Homogeneous effect", "Heterogeneous effect", "Confounding", 
                                          "Nonlinear effect", "Nonadditive effect"))
  
### Sensitivity
df_se <- all_weights %>%
  filter(scenario != "Null") %>%
  group_by(scenario, rho_X, method, seed) %>%
  summarise(se = sum(is_important & true_predictor) / sum(true_predictor))  %>%
  group_by(scenario, rho_X, method) %>%
  summarise(mean_se = mean(se, na.rm = TRUE))

ggplot(df_se[df_se$rho_X == 0.4, ], 
       aes(x = method, y = mean_se, fill = method)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ scenario) +                           
  labs(
    #title = "Mean sensitivity by scenario and method",
    x = "Method", y = "Sensitivity", fill = "Method") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

### Specificity
df_sp <- all_weights %>%
  filter(scenario != "Null") %>%
  group_by(scenario, rho_X, method, seed) %>%
  summarise(sp = sum(!is_important & !true_predictor) / sum(!true_predictor))  %>%
  group_by(scenario, rho_X, method) %>%
  summarise(mean_sp = mean(sp, na.rm = TRUE))

ggplot(df_sp[df_sp$rho_X == 0.4, ], 
       aes(x = method, y = mean_sp, fill = method)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ scenario) +                           
  labs(
    #title = "Mean sensitivity by scenario and method",
    x = "Method", y = "Specificity", fill = "Method") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())


### Plot weights
df_weights <- all_weights %>%
  group_by(scenario, rho_X, method, mix_name) %>%
  summarise(mean_weight = mean(abs(estimate), na.rm = TRUE)) %>%
  group_by(scenario, rho_X, method) %>%
  mutate(mean_weight = ifelse(method == "BKMR", mean_weight / sum(mean_weight), mean_weight)) %>%
  ungroup()

df_true_weights <- all_weights %>%
  group_by(scenario, rho_X, mix_name) %>%
  summarise(mean_weight = mean(true_weight, na.rm = TRUE)) %>%
  mutate(method = "truth")

df_weights <- rbind(df_weights, df_true_weights)
df_weights$method <- factor(df_weights$method, levels = c("truth", "BKMR", "BWS", "qgcomp", "WQS", "WQS2: pos"))

ggplot(df_weights[df_weights$rho_X == 0.7, ], 
       aes(x = method, y = mean_weight, fill = mix_name, alpha = factor(ifelse(scenario == "Null", "light", "dark")))) +
  geom_bar(stat = "identity", position = "stack") +  
  facet_wrap(~ scenario) +                           
  labs(
    #title = "Mean weights by scenario and method",
    x = "Method", y = "Weight", fill = "Exposure") +
  scale_alpha_manual(values = c("light" = 0.5, "dark" = 1)) + 
  guides(alpha = "none") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r eval=FALSE, include=FALSE}
library(scales)
for (scenario in all_results) {
p <- ggplot(scenario[[1]]$bkmr, aes(quantile, est, ymin = est - 1.96*sd, ymax = est + 1.96*sd)) +
  geom_pointrange() + 
  xlab("joint exposure quantile") +
  ylab("est")
print(p)
}

ggplot(all_bkmr, aes(x = quantile, y = est, ymin = est - 1.96 * sd, ymax = est + 1.96 * sd)) +
  geom_pointrange() + 
  facet_wrap(~ scenario + rho_X, ncol = 4) +
  labs(x = "joint exposure quantile", y = "est") +
  scale_x_continuous(labels = number_format(accuracy = 0.1)) +
  theme_minimal()
```

### AIC comparison
```{r}
# all_coef %>%
#   group_by(scenario, rho_X) %>%
#   summarise(unique_seed_count = n_distinct(seed), .groups = 'drop')

df_aic <- all_coef %>%
  filter(scenario %in% c("Nonlinear effect", "Nonadditive effect")) %>%
  group_by(scenario, rho_X, seed) %>%
  summarize(favor_nonlinear = AIC_qgcomp[1] > AIC_qgcomp_boot[1], .groups = "drop") %>%
  group_by(scenario, rho_X) %>%
  summarize(pct_favor_nonlinear = round(mean(favor_nonlinear), 4), .groups = "drop")
kable(df_aic, format = "markdown", caption = "Percentage of iterations favoring nonlinear qgcomp models")
```

